---
title: "Assignment 5"
author: "Cesar L. Espitia"
date: "3/4/2018"
output:
  html_document: default
  pdf_document: default
---

## Probabilities Assignment

Choose independently two numbers B and C at random from the interval [0, 1] with uniform density. Prove that B and C are proper probability distributions.   Note that the point (B,C) is then chosen at random in the unit square.  Find the probability:

### B + C < 0.5

```{r prob1-1}
set.seed(5) 
n=10000
B = runif(n, min = 0, max = 1)
set.seed(6)
C = runif(n, min = 0, max = 1)

sum((B+C)<0.5)/n

```
The probability of a value sum being less than 0.5 is 12.6%.

### B*C < 0.5

```{r prob1-2}
sum((B*C)<0.5)/n
```

The probability of a value product being less than 0.5 is 84%.

### |B - C| < 0.5

```{r prob1-3}
sum(abs(B-C)<0.5)/n
```

The probability of a value being less than 0.5 is 0.75%.


### max(B,C) < 0.5

```{r prob1-4}
count <- 0
for(i in 1:n){
  if(max(c(B[i],C[i])) < 0.5){
    count = count+1
  }
}
count/n
```

The probability of max(B,C) being less than 0.5 is 0.25%



### min(B,C) < 0.5

```{r prob1-5}
count <- 0
for(i in 1:n){
  if(min(c(B[i],C[i])) < 0.5){
    count = count+1
  }
}
count/n
```

The probability of min(B,C) being less than 0.5 is 0.75%
