---
title: "Data 605 - Assignment 3"
author: "Cesar Espitia"
date: "2/14/2018"
output:
  pdf_document: default
  html_document: default
---
# 1. Problem Set 1

### (1) What is the rank of Matrix A?

$$
A = \left|\begin{array}{cc} 
1 & 2 & 3 & 4\\
-1 & 0& 1 & 3 \\
0 & 1 & -2 & 1 \\
5 & 4 & -2 & -3
\end{array}\right|
$$ 

```{r setup, include=FALSE}
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3), nrow = 4)

qr(A)

B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow = 3)

qr(B)
```

The rank is 4.

### (2) Given $m\times n$ where $m>n$ what is max rank what is min rank (assuming non-zero)?

The max rank is n and the min rank is 1 assuming non-zero.

### (3) What is the rank of B?

$$
B = \left|\begin{array}{cc} 
1 & 2 & 1\\
3 & 6 & 3 \\
2 & 4 & 2
\end{array}\right|
$$ 

The rank is 1 because $r_{2} =2\times r_{1}$ and $r_~3~ _{3} =3\times r_{1}$.


# 2. Problem Set 2

Compute the eigenvalues and eigenvectors of the matrix A.

$$
A = \left|\begin{array}{cc} 
1 & 2 & 3 \\
0 & 4 & 5 \\
0 & 0 & 6
\end{array}\right|
$$ 

First we solve for 
$$
det \left|\begin{array}{cc} 
A - \lambda I 
\end{array}\right| 
$$
which equals:

$$
\left|\begin{array}{cc} 
1-\lambda & 2 & 3 \\
0 & 4-\lambda & 5 \\
0 & 0 & 6-\lambda
\end{array}\right|
$$ 

Reducing the matrix gives us:

$$
1-\lambda \left|\begin{array}{cc} 
4-\lambda & 5 \\
0 & 6-\lambda
\end{array}\right|
-2 \left|\begin{array}{cc} 
0 & 5 \\
0 & 6-\lambda
\end{array}\right|
+3 \left|\begin{array}{cc} 
0 & 4-\lambda \\
0 & 0
\end{array}\right|
$$ 

```{r two, include=FALSE}
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow = 3)

eigen(A)
```

This gives us when fully reduced $0 = (1-\lambda)(6-\lambda)(4-\lambda)$ which gives us the eigenvalues \lambda = 1,6,4 with eigenvectors 

$$
\left|\begin{array}{cc} 
0\\
0\\
1
\end{array}\right|
\left|\begin{array}{cc} 
0\\
0.37\\
-0.93
\end{array}\right|
\left|\begin{array}{cc} 
0.83\\
-0.55\\
0.06
\end{array}\right|
$$